---
title: "Machine Learning Project"
output: html_document
---

This is the R Markdown document for the Machine Learning Project.

**First we get the data, remove both columns that contain NA and the first two columns which are not relevant for prediction**
```{r, echo = TRUE}
library(caret)
set.seed(1234)
dtrain <- read.csv("pml-training.csv")
dtest <- read.csv("pml-testing.csv")
ctrain <- apply(is.na(dtrain), 2, sum) > dim(dtrain)[1]/2
ctest <- apply(is.na(dtest), 2, sum) > dim(dtest)[1]/2
cindex <- ctrain | ctest
cindex[1:2] <- TRUE

dtrain2 <- dtrain[, !cindex]
dtest2 <- dtest[, !cindex]
```

**Then we build training and validation sets and get our rf model**
```{r, echo = TRUE}
inTrain <- createDataPartition(y = dtrain2$classe, p = 0.7, 
                               list = FALSE)
dtrain3 <- dtrain2[inTrain,]
dvalid <- dtrain2[-inTrain,]
control1 <- trainControl(method = "cv")
mf1 <- train(classe ~ ., method = "rf", trControl = control1,
             data = dtrain3)
plot(mf1)
```

**Out-of-sample error, cross validation and prediction**

The expected out of sample error is 1 - accuracy and we can get accuracy from the following result:
```{r, echo = FALSE}
mf1
```
So we get the expected accuracy of
```{r, echo = FALSE}
accE <- mf1$results[match(mf1$bestTune, mf1$results[, 1]),2]
accE
```
and the error rate is
```{r, echo = FALSE}
1 - accE
```
We get the actual error rate using the validation set
```{r, echo = TRUE}
pred1 <- predict(mf1, dvalid)
accuA <- sum(pred1 == dvalid$classe) / length(dvalid$classe)  
err1 <- 1 - accuA
```
So the out-of-sample error is
```{r, echo = FALSE}
err1
```


**Finally we make a prediction on the test set. The correct classe values for the test set are known, so we can compare them with our prediction**

```{r, echo = TRUE}
answers <- c("B", "A", "B", "A", "A", "E", "D", "B", "A", "A",
             "B", "C", "B", "A", "E", "E", "A", "B", "B", "B")
pred2 <- predict(mf1, dtest2)
table(pred2, answers)
```

So we make a prediction on the test set with
`r sum(pred2 == answers) / length(answers) * 100`% accuracy.
